---
title: "Section10"
author: "Jax Lubkowitz"
format: 
  html:
    embed-resources: true
---

```{r}
library(tidyverse)
library(here)
library(broom)
theme_set(theme_minimal())

titanic_df <- read_csv(here("data/titanic.csv"),
                       col_types = list(Pclass = col_factor())) |>
  mutate(Pclass = fct_recode(Pclass,
                             "1st" = "1",
                             "2nd" = "2",
                             "3rd" = "3"))
```

```{r}
titanic_mod <- glm(Survived ~ Age,
                   data = titanic_df, family = "binomial")

survived <- titanic_df |> filter(Survived == 1)
dead <- titanic_df |> filter(Survived == 0)

titanic_mod |> tidy()

ggplot(data = titanic_df, aes(x = Age, y = Survived)) +
  geom_rug(data = survived, sides = "t") +
  geom_rug(data = dead, sides = "b")+
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  theme_minimal()
```

The model predicts that a one year increase in age is associated with a 0.00261 decrease in the log-odds of survival.

The model predicts that a one year increase in age is associated with a multiplicative change in the odds of survival of 0.997 times

# Exercise 1
```{r}
fare_mod <- glm(Survived ~ Fare, data = titanic_df, family = "binomial")
fare_mod |> tidy()

ggplot(data = titanic_df, aes(x = Fare, y = Survived)) +
  geom_jitter(height = 0.05) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  theme_minimal()
```




# Exercise 2
```{r}
sex_mod <- glm(Survived ~ Sex, data = titanic_df, family = "binomial")
sex_mod |> tidy()
```
The model predicts that males are associated with a -2.51 decrease in the log-odds of survival compared to females.


```{r}
titanic_large <- glm(Survived ~ Age + Sex + Pclass, data = titanic_df,
                     family = "binomial")
titanic_large |> tidy()
library(modelr)
grid <- titanic_df |>
  data_grid(
    Age = seq_range(Age, n = 10),
    Sex = c("female", "male"),
    Pclass = c("1st", "2nd", "3rd")
  ) 


# .fitted has predicted log odds which range from -inf to +inf
aug_surv <- augment(titanic_large, newdata = grid,
                    se_fit = TRUE)

ggplot(data = aug_surv, aes(x = Age, y = .fitted, color = Sex)) +
  geom_line() +
  facet_wrap(~ Pclass)

# Convert log odds to prob
aug_surv <- aug_surv |> mutate(pred_prob = exp(.fitted)/(1+exp(.fitted)))


ggplot(data = aug_surv, aes(x = ge, y = pred_prob, color = Sex)) +
  geom_line() +
  facet_wrap(~ Pclass)

```

# Adding interaction 
```{r}
titanic_int <- glm(Survived ~ Age + Sex + Pclass + Sex:Pclass, data = titanic_df,
                     family = "binomial")
aug_surv_int <- augment(titanic_int, newdata = grid,
                    se_fit = TRUE)
aug_surv_int <- aug_surv_int |> mutate(pred_prob = exp(.fitted)/(1+exp(.fitted)))

plot_df <- bind_rows(lst(aug_surv, aug_surv_int), .id = "model")

ggplot(data = plot_df, aes(x = Age, y = pred_prob)) +
  geom_line(aes(color = model)) +
  facet_grid(~ Pclass + Sex) +
  geom_rug(data = survived, sides = "t", aes(y =Survived)) +
  geom_rug(data = dead, sides = "b", aes(y = Survived ))
  

titanic_int |> glance()

titanic_large |> glance()
```
Interaction has better BIC and AIC (lower) so is better model. 
